---
title: Multinomial
author: kheagan
date: '2020-03-15'
slug: multinomial
categories: []
tags: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plotly)
library(dplyr)
```



# Multinomial notes

varibles and there catogries 

1. Heart attack : has had heart attach | han't had heart attack 
2. Colestral: over 260 | under 260



Assuming independence
$$p(A \ and \ B) = p(A)p(B) $$


$$p(A \ and \ B |X) = p(A | X)p(B |X)$$
where 

$$p(A | X) \propto l(A)p(A) $$

and simmarly for B,
$$p(B | X) \propto l(B)p(B) $$

We are doing this instead of 

$$p(A \ and \ B |X) \propto l(A  \ and \ B)p(A \ and \ B ) $$

I will try this with multinomial on simulated data 


I rewirte each cell i as $\pi_i$ and the count of observations that fall into that cell i as $x_i$

$$ f(x_1, x_2, x_3, x_4|\pi_1,\pi_3,\pi_3,\pi_4) = \frac{n!}{x_1!x_2!x_3!x_4!}\pi_1^{x_1} \pi_2^{x_2} \pi_3^{x_3} \pi_4^{x_4} $$

Note that the cells are independet. The varibles may or may not be depent, for example, the vrible  *has or has not heart attack* may depend on *cholsetral levels*. But this is taken care of when we account for all the possible comibnations of catogries. Thus by indpendece we can rewrite the posterior as

$$ p(\pi_1,\pi_2,\pi_3,\pi_4|x) \propto p(\pi_1| x)p(\pi_2| x)p(\pi_3| x) p(\pi_4| x) $$

,then for 
$$ p( \pi_1 | x) \propto l(\pi_1,\pi_2, \pi_3, \pi_4)p(\pi_1).$$




Where the liklihood can now be model by the binomail distribution, the prior by the beta distribution, and thus, the posterior by the beta distribution.
$$ p( \pi_i | x) \propto l(\pi_i)p(\pi_i) $$ 

The density of the data is binomial. 
$$ f(x| \pi_i) \propto  \pi_i^{x}(1-\pi_i)^{n-x} $$

The likihood of the paramters is
$$ l(\pi_i) \propto  \pi_i^{\sum x_i}(1-\pi_i)^{\sum (n-x_i)}$$

The prior is
$$ p(\pi_i) \propto \pi_i^{a-1}(1-\pi_i)^{b-1}$$

The posterior for each $\pi_i$ is 

$$p(\pi_i|x) \propto \pi_i^{a+\sum x_i-1}(1-\pi_i)^{b+\sum (n-x_i)-1} $$


####aside 
This is simular to the joint posterior distribution. The posterior of the joint paramters using a drikelt prior is 
$$ p(all \ \pi_i|x) \propto \prod_i\pi_i^{a_i+\sum x_i-1} $$
Which in our case would be 
$$ p(\pi_1,\pi_2,\pi_3,\pi_4|x) \propto \pi_1^{a_1+\sum x_1-1} \pi_2^{a_2+\sum x_2-1} \pi_3^{a_3+\sum x_3-1} \pi_4^{a_4+\sum x_4-1}$$
*But now we use gibbs samlping* and arive at the aformentioned
The posterior is 
$$ p(\pi_i|x) \propto \pi_i^{a+\sum x_i-1}(1-\pi_i)^{b+\sum (n-x_i)-1}$$

where $\pi_i|x$ is $beta(a+\sum x_i, b+\sum n-x_i)$


# An example



$$ P(A \ relisation| certain \ features)  = p( has \ heart \ attack \cap \ over \ 260)^{yes/no} p( has \ heart \ attack \cap \ under \ 260)^{yes/no}*$$
$$ p( hasn't \ heart \ attack \cap \ over \ 260)^{yes/no} p( hasn't \ heart \ attack \cap \ under \ 260)^{yes/no}$$


```{r}
HasHeartAttack <- c(51,41, 51+41)
HasNotHeartAttack <- c(992, 240, 992+240)
total <- c(1043,280, 1329)

data.frame(HasHeartAttack,HasNotHeartAttack,total, row.names = c('Cholestral over 260', 'Cholestral under 260', 'Total') )
```

My notation

total observations = 1329

1. $ \pi_1 $ = P( has had heart attack and cholestral over 260)
2. $ \pi_2 $ = P( has had heart attack and cholestral under 260)
3. $ \pi_3 $ = P( hasn't had heart attack and cholestral over 260)
4. $ \pi_4 $ = P( hasn't had heart attack and cholestral under 260)


1. $ \X_1 $ = count of indivdulas that has had heart attack and cholestral over 260
2. $ \X_2 $ = count of indivdulas that has had heart attack and cholestral under 260
3. $ \X_3 $ = count of indivdulas that hasn't had heart attack and cholestral over 260
4. $ \X_4 $ = count of indivdulas that hasn't had heart attack and cholestral under 260

The posterior is 
$$ p(\pi_i|x) \propto \pi_i^{a+\sum x_i-1}(1-\pi_i)^{b+\sum (n-x_i)-1}$$

```{r}
pi.possible.values <- seq(0,1,0.001)

a <- 0.5
b <- 0.5 

X_1 <- 51
X_2 <- 41
X_3 <- 992
X_4 <- 240

total <- 1329

pi_1 <- dbeta(pi.possible.values, a + X_1, b + total - X_1)
pi_2 <- dbeta(pi.possible.values, a + X_2, b + total - X_2)
pi_3 <- dbeta(pi.possible.values, a + X_3, b + total - X_3)
pi_4 <- dbeta(pi.possible.values, a + X_4, b + total - X_4)

sims.pi_1 <- rbeta(1000, a + X_1, b + total - X_1)
sims.pi_2 <- rbeta(1000, a + X_2, b + total - X_2)
sims.pi_3 <- rbeta(1000, a + X_3, b + total - X_3)
sims.pi_4 <- rbeta(1000, a + X_4, b + total - X_4)


plot_ly(x=pi.possible.values, y=log(pi_1), name = 'pi for has had heart attack and cholestral over 260', width =1000) %>% add_lines() %>%
  add_lines(x=pi.possible.values, y=log(pi_2), name = 'pi for has had heart attack and cholestral under 260') %>%
  add_lines(x=pi.possible.values, y=log(pi_3),  name = 'pi for has not had heart attack and cholestral over 260') %>%
  add_lines(x=pi.possible.values, y=log(pi_4),  name = 'pi for has not  had heart attack and cholestral under 260') %>%
  layout(title='log posterior for pi')


plot_ly(x=pi.possible.values, y=pi_1/max(pi_1), name = 'pi for has had heart attack and cholestral over 260', width =1000) %>% add_lines() %>%
  add_lines(x=pi.possible.values, y=pi_2/max(pi_2), name = 'pi for has had heart attack and cholestral under 260') %>%
  add_lines(x=pi.possible.values, y=pi_3/max(pi_3),  name = 'pi for has not had heart attack and cholestral over 260') %>%
  add_lines(x=pi.possible.values, y=pi_4/max(pi_4),  name = 'pi for has not  had heart attack and cholestral under 260') %>%
  layout(title='scaled posterior for pi')

plot_ly(x=sims.pi_1, name = 'pi for has had heart attack and cholestral over 260', alpha = 0.6, width =1000) %>% add_histogram() %>%
  add_histogram(x=sims.pi_2, name = 'pi for has had heart attack and cholestral under 260') %>%
  add_histogram(x=sims.pi_3,  name = 'pi for has not had heart attack and cholestral over 260') %>%
  add_histogram(x=sims.pi_4,  name = 'pi for has not  had heart attack and cholestral under 260') %>%
  layout(title='predictive density for pi', barmode='overlay')




```


```{r}
n.sims <- 1000
n.size <- 100
X_1.obs <- rbinom(n.sims,n.size,sims.pi_1)
X_2.obs <- rbinom(n.sims,n.size,sims.pi_2)
X_3.obs <- rbinom(n.sims,n.size,sims.pi_3)
X_4.obs <- rbinom(n.sims,n.size,sims.pi_4)



plot_ly(x=X_1.obs, name= 'observations for has had heart attack and cholestral over 260', width =1000) %>% add_histogram() %>%
   add_histogram(x= X_2.obs,  name = 'observations for has had heart attack and cholestral under 260')  %>%
   add_histogram(x= X_3.obs,  name = 'observations for has not had heart attack and cholestral over 260')  %>%
   add_histogram(x= X_4.obs, name= 'observations for has not  had heart attack and cholestral under 260') %>%
   layout(title='Simulations of new observations', barmode='overlay')


```



## Bring the freaq out the beyes 
### mean contgency table of simulated results 
```{r}

sims.has.heart.attack      <- c( mean(X_1.obs)       , mean(X_2.obs)        , mean(X_1.obs+X_2.obs))
sims.has.not.heart.attack  <- c( mean(X_3.obs)       , mean(X_4.obs)        , mean(X_3.obs+X_4.obs))
Total                      <- c(mean(X_1.obs+X_3.obs), mean(X_2.obs+X_4.obs), mean(X_1.obs+X_2.obs+X_3.obs+X_4.obs) )

df <- data.frame(sims.has.heart.attack,sims.has.not.heart.attack, Total, row.names = c('cholestral over 260', 'cholestral under 260','Total'))
colnames(df) <- c('has had heart attack','has not had heart attack','Total')

df
#total = sum(X_1.obs+X_2.obs +X_3.obs+X_4.obs )
```




If two events are independent then..
$$p(A \cap B) = p(A)p(B)$$
We can construct a hypothesis test. The null hypothesis stats that *more or less*
$$ 0 = p(A \cap B) - p(A)p(B)$$
We will test this hypothesis with our simulated observations
```{r}
checking.indepedece.X.1.obs <- (X_1.obs/(X_1.obs+X_2.obs+X_3.obs+X_4.obs))  - ((X_1.obs/(X_1.obs+X_3.obs))*(X_1.obs/(X_1.obs+X_2.obs)))
checking.indepedece.X.2.obs <- (X_2.obs/(X_1.obs+X_2.obs+X_3.obs+X_4.obs))  - ((X_2.obs/(X_2.obs+X_4.obs))*(X_1.obs/(X_1.obs+X_2.obs)))
checking.indepedece.X.3.obs <- (X_3.obs/(X_1.obs+X_2.obs+X_3.obs+X_4.obs))  - ((X_3.obs/(X_1.obs+X_3.obs))*(X_1.obs/(X_3.obs+X_4.obs)))
checking.indepedece.X.4.obs <- (X_4.obs/(X_1.obs+X_2.obs+X_3.obs+X_4.obs))  - ((X_4.obs/(X_2.obs+X_4.obs))*(X_1.obs/(X_3.obs+X_4.obs)))



fig1 <- plot_ly(x=checking.indepedece.X.1.obs, name='observations for has had heart attack and cholestral over 260', width = 1200) %>%
  add_histogram() %>%
  layout(title='Testing if aproximatly 0 = P(AB) -P(A)P(B) ')
fig2 <- plot_ly(x=checking.indepedece.X.2.obs, name='observations for has had heart attack and cholestral under 260') %>% add_histogram()
fig3 <- plot_ly(x=checking.indepedece.X.3.obs,  name = 'observations for has not had heart attack and cholestral over 260') %>% add_histogram()
fig4 <- plot_ly(x=checking.indepedece.X.4.obs, name= 'observations for has not  had heart attack and cholestral under 260') %>% add_histogram()
subplot(fig1, fig2, fig3, fig4, nrows = 2, which_layout = 1)

```

From the above we can reject the null hypothesis for has not had heart attack and cholestral over 26. And conclude there is a relationship between these events. By that logic the statment 

$$p(ABCD) = p(A)p(B)p(C)p(B)$$
is false. And conclude that there is a relationship between cholestral and heart attack.


# latent class notes

Consider if pateints where experiencing symptoms A, B and C.

1. ) A : yes/ no
2. ) B : yes/ no
3. ) C : yes/ no

And we wanted to model if they have disease X, Y and Z




$$  P(Has \ disease \ z_i)   =  p(A \cap B \cap C| z_i)^{yes/no}*p(A \cap B \cap \not C| z_i)^{yes/no}*p(A \cap \not B \cap C| z_i)^{yes/no}*p(A \cap \not B \cap \not C| z_i)^{yes/no}* $$

$$p(\not A \cap B \cap C| z_i)^{yes/no}*p( \not A \cap B \cap \not C| z_i)^{yes/no}*p(\not A \cap \not B \cap C| z_i)^{yes/no}*p(A \cap \not B \cap \not C| z_i)^{yes/no}$$

 


That is a huge model for very simple data. This is where latent classes come in. So instead we model, because of local independece...

$$  P(Has \ disease \ z_i)   =  p(A| zi)^{yes/no}* p(B| zi)^{yes/no}* p(C| zi)^{yes/no}$$


This is how we will get the final results

$$   P(z_1, z_2,z_3 | symptoms \ A,B,C ) \propto P(Has \ disease \ z_1)  P(Has \ disease \ z_2) P(Has \ disease \ z_3)   $$ 


### simulation
```{r}
n.sims <- 1000

#these would be multinomial for more than 2 catogries


#10 pateints come into the hospital 

rows <- matrix(NA, nrow = n.sims, ncol = 4)

for (i in 1:n.sims) {
  Disease <- t(rmultinom(1, 1, c(0.5,0.2,0.3)))[1,]
  X <-  Disease[1]
  Y <-  Disease[2]
  
  D <- NA
  symptoms <- c(NA,NA,NA)
  
  
  if(X == 1){
    symptoms <- t(rmultinom(1, 1, c(0.1,0.8,0.1)))[1,]
    D <- 'X'
    
  } else if(Y == 1){
    
    symptoms <- t(rmultinom(1, 1, c(0,0.2,0.8)))[1,]
    D <- 'Y'  
    
  }else{
    symptoms <- t(rmultinom(1, 1, c(0.9,0.05,0.05)))[1,]
    D <- 'Z'  
  }
  
 rows[i,] <-  c(symptoms, D)
}


df <- as.data.frame(rows)
colnames(df) <- c('A','B','C','Disease')
head(df)

summary(df)
```


So forgeting how i made the table. Let's start 

```{r}

X <-df[df$Disease == 'X',]
Y <-df[df$Disease == 'Y',]
Z <-df[df$Disease == 'Z',]

table(df)
#plot_ly() %>% add_pie(data = X[], values = ~n)

```





Let's consider disease X

```{r}
head(X)
```



*But now we use gibbs samlping* and arive at the aformentioned
The posterior probablity $\pi $ for each symptom is  
$$ p(\pi_i|x) \propto \pi_i^{a+\sum x_i-1}(1-\pi_i)^{b+\sum (n-x_i)-1}$$

```{r}
A <-table(X$A)[2]
B <- table(X$B)[2]
C <- table(X$C)[2]

a <- 0.5
b <- 0.5

sum.A <- A
sum.B <- B
sum.C <- C

x<- c(sum.A, sum.B, sum.C)

Total <- sum.A + sum.B + sum.C

param <- seq(0,1,0.001)
pi.A <- dbeta(param, a+sum.A, b + Total-sum.A , log = TRUE)
pi.B <- dbeta(param, a+sum.B, b + Total-sum.B , log = TRUE)
pi.C <- dbeta(param, a+sum.C, b + Total-sum.C , log = TRUE)

n.sims <- 1000
X.sims.pi.A <- rbeta(n.sims, a+sum.A, b + Total-sum.A)
X.sims.pi.B <- rbeta(n.sims, a+sum.B, b + Total-sum.B)
X.sims.pi.C <- rbeta(n.sims, a+sum.C, b + Total-sum.C)


fig1 <- plot_ly(x=param, y=pi.A, name='log post pi symptoms A', width = 1000) %>% add_lines() %>% layout(title='Disease X',  barmode='overlay') %>%
add_lines(x=param, y=pi.B, name='log post pi symptoms B') %>% 
add_lines(x=param, y=pi.C, name='log post pi symptoms C')

fig2 <- plot_ly(x=param, y=exp(pi.A-max(pi.A)), name='scaled pi post symptoms A') %>% add_lines() %>%
add_lines(x=param, y=exp(pi.B-max(pi.B)), name='scaled post pi symptoms B') %>%
add_lines(x=param, y=exp(pi.C-max(pi.C)), name='scaled post pi symptoms C') 

fig3 <- plot_ly(x=X.sims.pi.A, name='sims pi symptoms A',alpha = 0.7) %>% add_histogram() %>% 
add_histogram(x=X.sims.pi.B, name='sims pi symptoms B') %>%
add_histogram(x=X.sims.pi.C, name='sims pi symptoms C')


subplot(fig1,fig2, fig3, which_layout = 1)
```



```{r}
A <-table(Y$A)[2]
B <- table(Y$B)[2]
C <- table(Y$C)[2]
 
sum.A <- A
sum.B <- B
sum.C <- C

y <- c(sum.A, sum.B, sum.C)
 

a <- 0.5
b <- 0.5


Total <- sum.A + sum.B + sum.C

param <- seq(0,1,0.001)
pi.A <- dbeta(param, a+sum.A, b + Total-sum.A , log = TRUE)
pi.B <- dbeta(param, a+sum.B, b + Total-sum.B , log = TRUE)
pi.C <- dbeta(param, a+sum.C, b + Total-sum.C , log = TRUE)

n.sims <- 1000
Y.sims.pi.A <- rbeta(n.sims, a+sum.A, b + Total-sum.A)
Y.sims.pi.B <- rbeta(n.sims, a+sum.B, b + Total-sum.B)
Y.sims.pi.C <- rbeta(n.sims, a+sum.C, b + Total-sum.C)


fig1 <- plot_ly(x=param, y=pi.A, name='log post pi symptoms A', width = 1000) %>% add_lines() %>% layout(title='Disease Y',  barmode='overlay') %>%
add_lines(x=param, y=pi.B, name='log post pi symptoms B') %>% 
add_lines(x=param, y=pi.C, name='log post pi symptoms C')

fig2 <- plot_ly(x=param, y=exp(pi.A-max(pi.A)), name='scaled pi post symptoms A') %>% add_lines() %>%
add_lines(x=param, y=exp(pi.B-max(pi.B)), name='scaled post pi symptoms B') %>%
add_lines(x=param, y=exp(pi.C-max(pi.C)), name='scaled post pi symptoms C') 

fig3 <- plot_ly(x=Y.sims.pi.A, name='sims pi symptoms A',alpha = 0.7) %>% add_histogram() %>% 
add_histogram(x=Y.sims.pi.B, name='sims pi symptoms B') %>%
add_histogram(x=Y.sims.pi.C, name='sims pi symptoms C')


subplot(fig1,fig2, fig3, which_layout = 1)
```


```{r}
A <-table(Z$A)[2]
B <- table(Z$B)[2]
C <- table(Z$C)[2]
 
sum.A <- A
sum.B <- B
sum.C <- C


a <- 0.5
b <- 0.5
z <- c(sum.A, sum.B, sum.C)

Total <- sum.A + sum.B + sum.C


param <- seq(0,1,0.001)
pi.A <- dbeta(param, a+sum.A, b + Total-sum.A , log = TRUE)
pi.B <- dbeta(param, a+sum.B, b + Total-sum.B , log = TRUE)
pi.C <- dbeta(param, a+sum.C, b + Total-sum.C , log = TRUE)

n.sims <- 1000
Z.sims.pi.A <- rbeta(n.sims, a+sum.A, b + Total-sum.A)
Z.sims.pi.B <- rbeta(n.sims, a+sum.B, b + Total-sum.B)
Z.sims.pi.C <- rbeta(n.sims, a+sum.C, b + Total-sum.C)


fig1 <- plot_ly(x=param, y=pi.A, name='log post pi symptoms A', width = 1000) %>% add_lines() %>% layout(title='Disease Z',  barmode='overlay') %>%
add_lines(x=param, y=pi.B, name='log post pi symptoms B') %>% 
add_lines(x=param, y=pi.C, name='log post pi symptoms C')

fig2 <- plot_ly(x=param, y=exp(pi.A-max(pi.A)), name='scaled pi post symptoms A') %>% add_lines() %>%
add_lines(x=param, y=exp(pi.B-max(pi.B)), name='scaled post pi symptoms B') %>%
add_lines(x=param, y=exp(pi.C-max(pi.C)), name='scaled post pi symptoms C') 

fig3 <- plot_ly(x=Z.sims.pi.A, name='sims pi symptoms A',alpha = 0.7) %>% add_histogram() %>% 
add_histogram(x=Z.sims.pi.B, name='sims pi symptoms B') %>%
add_histogram(x=Z.sims.pi.C, name='sims pi symptoms C')


subplot(fig1,fig2, fig3, which_layout = 1)
```



```{r}

a <- 0.5
b <- 0.5

sum.X <- length(X$Disease)
sum.Y <- length(Y$Disease)
sum.Z <- length(Z$Disease)

Total <- sum.X + sum.Y + sum.Z

param <- seq(0,1,0.001)
pi.X <- dbeta(param, a+sum.X, b + Total-sum.X , log = TRUE)
pi.Y <- dbeta(param, a+sum.Y, b + Total-sum.Y , log = TRUE)
pi.Z <- dbeta(param, a+sum.Z, b + Total-sum.Z , log = TRUE)

n.sims <- 1000
sims.pi.X <- rbeta(n.sims, a+sum.X, b + Total-sum.X)
sims.pi.Y <- rbeta(n.sims, a+sum.Y, b + Total-sum.Y)
sims.pi.Z <- rbeta(n.sims, a+sum.Z, b + Total-sum.Z)


fig1 <- plot_ly(x=param, y=pi.X, name='log post pi n cases dieases X', width = 1000) %>% add_lines() %>% layout(title='Propotion of patients with a certian disease',  barmode='overlay') %>%
add_lines(x=param, y=pi.Y, name='log post pi n cases dieases Y') %>% 
add_lines(x=param, y=pi.Z, name='log post pi n cases dieases Z')

fig2 <- plot_ly(x=param, y=exp(pi.X-max(pi.X)), name='scaled pi post n cases dieases X') %>% add_lines() %>%
add_lines(x=param, y=exp(pi.Y-max(pi.Y)), name='scaled post n cases pi dieases Y') %>%
add_lines(x=param, y=exp(pi.Z-max(pi.Z)), name='scaled post n cases pi dieases Z') 

fig3 <- plot_ly(x=sims.pi.X, name='sims pi n cases dieases X',alpha = 0.7) %>% add_histogram() %>% 
add_histogram(x=sims.pi.Y, name='sims pi n cases dieases Y') %>%
add_histogram(x=sims.pi.Z, name='sims pi n cases dieases Z')


subplot(fig1,fig2, fig3, which_layout = 1)



```



# Before we get the freq out of bayes 

### a quick summary of what we observed

A pivot table of disease and sympotoms
```{r}
observ.table <- data.frame(x,y,z, row.names = c('A', 'B','C'))
observ.table 

plot_ly() %>% add_heatmap(x=c('X','Y','Z'), y=c('A','B','C'), z=as.matrix(observ.table)) %>% layout(title='heatmap of observations')



```









```{r}
X.sim.obs <- rbinom(1000, 100, sims.pi.X )
Y.sim.obs <- rbinom(1000, 100, sims.pi.Y )
Z.sim.obs <- rbinom(1000, 100, sims.pi.Z )

plot_ly(alpha = 0.7, width=1100) %>% add_histogram(x= X.sim.obs, name='Number of pateinets with disease X')  %>% 
  add_histogram(x= Y.sim.obs, name='Number of pateinets with disease Y')  %>% 
  add_histogram(x= Z.sim.obs, name='Number of pateinets with disease Z') %>% 
  layout(title='histogram of simulated number of patents with a certain disease(out of 100 patents)', barmode='overlay')

```


## showing how the diffrent n simulate number of paentes (out of 100) who have disease x, affect simulations of symptom A given disease X 

```{r}

plot(ecdf(X.sim.obs))

samples.X.A <-t(as.matrix(sapply(sort(X.sim.obs), function(n){
  
  bin <- rbinom(1000, n, X.sims.pi.A) 
  table(cut(bin, 50))
  } )))

 

plot_ly() %>% add_heatmap(z=samples.X.A)




```













simulations for amount of cases of each disease 
```{r}

samples.X.A <- sapply(X.sim.obs, function(n){
  A.obs <- rbinom(100, n, X.sims.pi.A)
  mean(A.obs)
  })
samples.X.B <- sapply(X.sim.obs, function(n){
  B.obs <- rbinom(100, n, X.sims.pi.B)
  mean(B.obs)
  })
samples.X.C <- sapply(X.sim.obs, function(n){
  C.obs <- rbinom(100, n, X.sims.pi.C)
  mean(C.obs)
  })

samples.Y.A <- sapply(Y.sim.obs, function(n){
  A.obs <- rbinom(100, n, Y.sims.pi.A)
  mean(A.obs)
  })
samples.Y.B <- sapply(Y.sim.obs, function(n){
  B.obs <- rbinom(100, n, Y.sims.pi.B)
  mean(B.obs)
  })
samples.Y.C <- sapply(Y.sim.obs, function(n){
  C.obs <- rbinom(100, n, Y.sims.pi.C)
  mean(C.obs)
  })

samples.Z.A <- sapply(Z.sim.obs, function(n){
  A.obs <- rbinom(100, n, Z.sims.pi.A)
  mean(A.obs)
  })
samples.Z.B <- sapply(Z.sim.obs, function(n){
  B.obs <- rbinom(100, n, Z.sims.pi.B)
  mean(B.obs)
  })
samples.Z.C <- sapply(Z.sim.obs, function(n){
  C.obs <- rbinom(100, n, Z.sims.pi.C)
  mean(C.obs)
  })







fig1 <- plot_ly(x=samples.X.A, name='symptom A', color = 'blue', width = 1000) %>% add_histogram() %>%
add_histogram(x=samples.X.B, name='symptom B', color='red') %>% 
add_histogram(x=samples.X.C, name='Symptom C', color='green') %>%
  layout(title='Simulations for occurances a diseases symptoms( out of 100)', barmode='overlay', xaxis=list(title='disease X'))


fig2 <- plot_ly(x=samples.Y.A, name='symptom A',  color = 'blue') %>% add_histogram() %>%
add_histogram(x=samples.Y.B, name='symptom B', color='red') %>% 
add_histogram(x=samples.Y.C, name='Symptom C',  color='green') %>%
  layout(title='For disease Y', barmode='overlay', xaxis=list(title='disease Y'))

fig3 <- plot_ly(x=samples.Z.A, name='symptom A',  color = 'blue') %>% add_histogram() %>%
add_histogram(x=samples.Z.B, name='symptom B', color='red') %>% 
add_histogram(x=samples.Z.C, name='Symptom C', color='green') %>%
  layout(title='For disease Z', barmode='overlay', xaxis=list(title='disease Z'))

subplot(fig1, fig2, fig3, titleX = TRUE, which_layout = 1)


```


